Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping

### 2017-08-30 stp/paper_aaai/main.tex
Made couple of sentences shorter, made tables for the features intead of enums

\usepackage{color}
\usepackage{xcolor}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algrenewcommand\algorithmiccomment[2][\footnotesize]{{#1\hfill\(\triangleright\)
 #2}} %\normalsize
\usepackage{amssymb}
\usepackage{textcomp}

@@

\usepackage{tikz}
\tikzset{every picture/.style={font issue=\small},
         font issue/.style={execute at begin picture={#1\selectfont}}
        }
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{arrows,shapes,decorations.pathmorphing}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{mindmap,backgrounds}
\tikzstyle{class}=[draw,green,ellipse,align=center,fill=green!25,text=black]
\tikzstyle{data}=[draw,rectangle,align=center,black!70]
\tikzstyle{att}=[rectangle,align=center,fill=brown!70!red]
\tikzstyle{hasaA} = 
[draw,densely dashed,-,black!70,decoration={snake,segment 
length=10,amplitude=0.3,post length=4}]
\tikzstyle{hasaC} = [>=stealth',draw,thick,-, black]
\tikzstyle{isa} = [>=o,draw,densely dotted,very thick,black!90]

@@

In this paper we use Machine Learning techniques (ML) to learn mapping rules from previously mapped instances.
To this end we formulate the Relational-To-Ontology Mapping Problem (\relonto{}) as a Minimum Cost Steiner Tree Problem (STP) with side constraints \cite{deuna2016steiner}.

Firstly, we build an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. 
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming (CP) to 
find a minimum cost Steiner Tree in the graph.
The goal is to assign the costs on the edges in such a way that the resulting 
Steiner Tree is a valid and coherent semantic model for the new source. 
The contributions of this work are:
\begin{itemize}
	\item A novel modeling framework for the \relonto{} problem which 
	seamlessly incorporates machine learning and constraint programming.
	\item An efficient approach to handle attributes which cannot be matched to 
	the ontology, through a new semantic labeling model and a new STP 
	model with additional nodes.
	\item A flexible and extensible approach to the problem, which is easy to 
	reuse and adapt to specific situations.
\end{itemize}

@@

We formulate the problem of semantic labeling as a multi-class classification
problem.
The known source descriptions $S_T$ provide us the training sample.
We compute a feature vector for each attribute in a data source and associate 
the known semantic type with the corresponding feature vector.
The feature vector, among others, includes characteristics
such as a number 
of whitespaces and other special characters, statistics of 
values in the column 
(e.g, mean/max/min string length and numeric statistics), etc. 
A full list can be found in appendix \ddg{Apendix!}
\npr{If I remove entropy descr, we are perfectly within page limit}
%\ignore{
One of the important features characterising information content of an attribute is Shannon's entropy of the attribute's concatenated rows.
Shannon's entropy (or information entropy~\cite{Manning:Introduction}) of a string $X$ is defined as
$H(X) = -\sum_{i}{p_i \log_{2}p_i},$ where $p_i$ is the probability of a character, whose index in character vocabulary is $i$, to appear in $X$, and the summation ranges over all
characters in the vocabulary of all possible characters.
To evaluate $p_i$ in Shannon's entropy, we evaluate normalized character frequency distribution \emph{chardist} of an attribute, as character counts in concatenated rows of the attribute, normalized by the total length of the concatenated rows.
The vocabulary of all characters consists of 100 printable characters (including $\backslash$n).
Finally, we add the 100-dimensional vector of $p_i$ to the attribute feature vector.
%}
We also compute a set of features based on similarity metrics inspired by state-of-the-art works~\cite{Pham:semantic} or~\cite{Ritze:matching}.
Among others, we compute mean cosine similarity for character distributions of attribute values and string similarity metrics for attribute names.
%To extract features from attribute names, we compute string similarity metrics: minimum edit distance, two WordNet based similarity measures such as JCN~\cite{Jiang:Semantic} and LIN ~\cite{Lin:Information}, and $k$-nearest neighbors using Needle-Wunsch distance~\cite{Needleman:General}.
%The minimum edit distance between two strings $s_1$ and $s_2$ is the minimum number of edit operations, such as insertion, deletion, substitution, which are required to transform one string into another~\cite{Manning:Introduction}.
%We compute the similarity between attribute name and all class instances in the training data.
%The number of thus extracted features depends on the number of semantic labels in the training data.
We train a random forest on the obtained sample.

@@

Note that, although the alignment graph is directed, the semantic models are 
not necessarily rooted directed trees (e.g. Fig. \ref{FIG:sem}), therefore the 
direction of the edges is ignored for the STP model (and are all treated as 
undirected edges). 
\npr{This is not true: semantic models are directed. can we soften the statement here? direction of links can change semantics}

@@

Briefly, the way a CP solver works is by assigning a value to each variable of 
$\boldsymbol{v}$ in turn. At each iteration it will check whether any 
constraint in $C$ is violated, in which case it will backtrack to change its 
last decision. \ignore{The choice of the order of the variables to be assigned, 
and the 
choice of the value to be assigned to each variable is called the \emph{search 
strategy}.}
Since only using a backtracking algorithm would have a prohibitive cost, CP 
uses a combination of search and \emph{propagation}. The latter is achieved by  
removing inconsistent values 
from the domains of variables. That way, values that cannot appear in a 
solution, given the current set of decisions, are never tried by the search. 
\emph{Global constraints} are higher order constraints that enforce a 
complex 
constraint over a set of variables. These are typically implemented with 
specialized algorithms for these 
constraints and perform substantially better than a decomposition approaches
\cite{regin1994filtering,beldiceanu2012global}.

\subsection{\relonto{} in Constraint Programming}
In order to model this problem we used the \minizinc{} language 
\cite{minizinc}, and the 
\chuffed{} solver \cite{chu2011improving} because it has a global constraint 
implemented for Steiner 
Tree Problems \cite{deuna2016steiner}.

@@

\newpage

Below are the description of features used by \serene{} semantic model.
Note that features are extracted from the name and values of each attribute in the data source.
Thus, all the data from a single column is reduced to some feature
vector of the concatenation of values computed by the extractors below.

We refer to a single value from the column as an entry.
E.g.

@@

From the table above, John is an entry of the column Person.

The following feature extractors produce a scalar value:
\begin{enumerate}
\item num-unique-vals      : A simple count of unique entries of a column
\item prop-unique-vals     : The proportion of unique entries of a column. 
\item prop-missing-vals    : The proportion of all missing/empty entries from all entries in a column.
\item ratio-alpha-chars    : The average of the proportion of alphabetic characters from each entry.
\item prop-numerical-chars : The average of the proportion of numeric characters from each entry.
\item prop-whitespace-chars       : The average of the proportion of whitespace characters from each entry.
\item prop-entries-with-at-sign   : The proportion of entries with an '@' sign.
\item prop-entries-with-hyphen    : The proportion of entries with a '-'.
\item prop-range-format           : The proportion of entries which follow some numerical range format: ([0-9]+)-([0-9]+).
\item is-discrete                 : A binary indicator if the column entries are discrete.
\item entropy-for-discrete-values : The entropy of the entries of a column.
\item shannon-entropy             : Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'
\end{enumerate}

The following feature extractors produce a vector value.
Some of these are computed together to reduce computation time.
Others are values that are computed for all classes.

@@

\end{enumerate}

>>>>

\usepackage{color,colortbl}
\usepackage{xcolor}
\definecolor{Gray}{gray}{0.96}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algrenewcommand\algorithmiccomment[2][\footnotesize]{{#1\hfill\(\triangleright\)
 #2}} %\normalsize
\usepackage{amssymb}
\usepackage{textcomp}

@@

\usepackage{supertabular}

\usepackage{tikz}
\tikzset{every picture/.style={font issue=\small},
         font issue/.style={execute at begin picture={#1\selectfont}}
        }
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{arrows,shapes,decorations.pathmorphing}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{mindmap,backgrounds}
\tikzstyle{class}=[draw,green,ellipse,align=center,fill=green!25,text=black,inner sep=1]
\tikzstyle{data}=[draw,rectangle,align=center,black!70,inner sep=2]
\tikzstyle{att}=[rectangle,align=center,fill=brown!70!red]
\tikzstyle{hasaA} = 
[draw,densely dashed,-,black!70,decoration={snake,segment 
length=10,amplitude=0.3,post length=4}]
\tikzstyle{hasaC} = [>=stealth',draw,thick,-, black]
\tikzstyle{isa} = [>=o,draw,densely dotted,very thick,black!90]

@@

In this paper we use Machine Learning techniques (ML) to learn mapping rules from previously mapped instances.
To this end we formulate the Relational-To-Ontology Mapping Problem (\relonto{}) as a Minimum Steiner Tree Problem (STP) with side constraints \cite{deuna2016steiner}.

Firstly, we build an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. 
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming (CP) to 
find a minimum cost Steiner Tree in the graph.
The goal is to assign the costs on the edges in such a way that the resulting 
Steiner Tree is a valid and coherent semantic model for the new source. 
Our contributions are:
\begin{itemize}
	\item A novel modeling framework for the \relonto{} problem which 
	seamlessly incorporates machine learning and constraint programming.
	\item An efficient approach to handle attributes which cannot be matched to 
	the ontology, through a new semantic labeling model and a new STP 
	model with additional nodes.
	\item A flexible and extensible approach to the problem, which is easy to 
	reuse and adapt to specific situations.
\end{itemize}

@@

We formulate the problem of semantic labeling as a multi-class classification
problem.
The known source descriptions $S_T$ provide us the training sample.
We compute a feature vector for each attribute in a data source and associate 
the known semantic type with the corresponding feature vector.
The feature vector, among others, includes characteristics
such as a number 
of whitespaces and other special characters, statistics of 
values in the column 
(e.g, mean/max/min string length and numeric statistics), etc. 
A full list can be found in Appendix \ref{Ann:SemLab}.
\npr{If I remove entropy descr, we are perfectly within page limit}
%\ignore{
One of the important features characterising information content of an attribute is Shannon's entropy of the attribute's concatenated rows.
Shannon's entropy (or information entropy~\cite{Manning:Introduction}) of a string $X$ is defined as
$H(X) = -\sum_{i}{p_i \log_{2}p_i},$ where $p_i$ is the probability of a character, whose index in character vocabulary is $i$, to appear in $X$, and the summation ranges over all
characters in the vocabulary of all possible characters.
To evaluate $p_i$ in Shannon's entropy, we evaluate normalized character frequency distribution \emph{chardist} of an attribute, as character counts in concatenated rows of the attribute, normalized by the total length of the concatenated rows.
The vocabulary of all characters consists of 100 printable characters (including $\backslash$n).
Finally, we add the 100-dimensional vector of $p_i$ to the attribute feature vector.
%}
We also compute a set of features based on similarity metrics inspired by state-of-the-art works~\cite{Pham:semantic} or~\cite{Ritze:matching}.
Among others, we compute mean cosine similarity for character distributions of attribute values and string similarity metrics for attribute names.
%To extract features from attribute names, we compute string similarity metrics: minimum edit distance, two WordNet based similarity measures such as JCN~\cite{Jiang:Semantic} and LIN ~\cite{Lin:Information}, and $k$-nearest neighbors using Needle-Wunsch distance~\cite{Needleman:General}.
%The minimum edit distance between two strings $s_1$ and $s_2$ is the minimum number of edit operations, such as insertion, deletion, substitution, which are required to transform one string into another~\cite{Manning:Introduction}.
%We compute the similarity between attribute name and all class instances in the training data.
%The number of thus extracted features depends on the number of semantic labels in the training data.
We train a random forest on the obtained sample.

@@

Note that, although the alignment graph is directed, the semantic models are 
not necessarily rooted graphs (e.g. Fig. \ref{FIG:sem}), therefore the 
direction of the edges is ignored for the STP model (and are all treated as 
undirected edges).

@@

Briefly, the way a CP solver works is by assigning a value to each variable of 
$\boldsymbol{v}$ in turn. At each iteration it will check whether any 
constraint in $C$ is violated, in which case it will backtrack to change its 
last decision. \ignore{The choice of the order of the variables to be assigned, 
and the 
choice of the value to be assigned to each variable is called the \emph{search 
strategy}.}
CP uses a combination of search and \emph{propagation}. The latter is achieved by  
removing inconsistent values 
from the domains of variables. That way, values that cannot appear in a 
solution, given the current set of decisions, are never tried by the search. 
\emph{Global constraints} are higher order constraints that enforce a 
complex 
constraint over a set of variables. These are typically implemented with 
specialized algorithms that perform substantially better than decomposition approaches
\cite{regin1994filtering,beldiceanu2012global}.

\subsection{\relonto{} in Constraint Programming}
In order to model this problem we used the \minizinc{} language 
\cite{minizinc}, and the 
\chuffed{} solver \cite{chu2011improving} because it has a global constraint 
implemented for Steiner 
Tree Problems \cite{deuna2016steiner}. The model can be found in Appendix \ref{Ann:MZ}.

@@

\newpage
\section*{Appendices}
\renewcommand{\thesubsection}{\Alph{subsection}}

\subsection{Features for Semantic Labeling\label{Ann:SemLab}}

Below are the description of features used by \serene{} semantic model.
Note that features are extracted from the name and values of each attribute in the data source.
Thus, all the data from a single column is reduced to some feature
vector of the concatenation of values computed by the extractors below.
We refer to a single value from the column as an entry. For example, in the table below, John is an entry of the column Person.

@@

\begin{table}[!ht]
	\centering
	\caption{Feature extractors that produce a scalar value}
	\begin{tabular}{p{3cm}p{4.5cm}}
	Feature name & Description \\ \hline
	num-unique-vals & A simple count of unique entries of a column\\
	\rowcolor{Gray}
	prop-unique-vals & The proportion of unique entries of a column. \\
	\rowcolor{Gray}
	prop-missing-vals & The proportion of all missing/empty entries from all entries in a column.\\
	ratio-alpha-chars & The average of the proportion of alphabetic characters from each entry.\\
	\rowcolor{Gray}
	prop-numerical-chars & The average of the proportion of numeric characters from each entry.\\
	prop-whitespace-chars    & The average of the proportion of whitespace characters from each entry.\\
	\rowcolor{Gray}
	prop-entries-with-at-sign& The proportion of entries with an '@' sign. \\
	prop-entries-with-hyphen & The proportion of entries with a '-'. \\
	\rowcolor{Gray}
	prop-range-format        & The proportion of entries which follow some \\ numerical range format &  ([0-9]+)-([0-9]+) \\
	\rowcolor{Gray}
	is-discrete              & A binary indicator if the column entries are discrete. \\
	entropy-for-discrete-values & The entropy of the entries of a column.\\
	\rowcolor{Gray}	
	shannon-entropy          & Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'\\
	\end{tabular}
\end{table}
\ignore{
The following feature extractors produce a scalar value:
\begin{enumerate}
\item num-unique-vals      : A simple count of unique entries of a column
\item prop-unique-vals     : The proportion of unique entries of a column. 
\item prop-missing-vals    : The proportion of all missing/empty entries from all entries in a column.
\item ratio-alpha-chars    : The average of the proportion of alphabetic characters from each entry.
\item prop-numerical-chars : The average of the proportion of numeric characters from each entry.
\item prop-whitespace-chars       : The average of the proportion of whitespace characters from each entry.
\item prop-entries-with-at-sign   : The proportion of entries with an '@' sign.
\item prop-entries-with-hyphen    : The proportion of entries with a '-'.
\item prop-range-format           : The proportion of entries which follow some numerical range format: ([0-9]+)-([0-9]+).
\item is-discrete                 : A binary indicator if the column entries are discrete.
\item entropy-for-discrete-values : The entropy of the entries of a column.
\item shannon-entropy             : Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'
\end{enumerate}
}

\begin{table}[!ht]
	\centering
	\caption{Feature extractors that produce a vector value. Some of these are computed together to reduce computation time.
	Others are values that are computed for all classes.}

	\begin{tabular}{p{2.5cm}p{5cm}}
	Feature name & Description \\ \hline
inferred-data-type      & Produces a vector which contains the inferred data type.  
Only a single element is non-zero in the vector (i.e. 1-of-K format).  
The set of inferred types are in float, integer, long, boolean, date, time, datetime, string. \\
	\rowcolor{Gray}
char-dist-features      & Produces a vector for character distribution of column values;
considered characters are printable characters
%                        0123456789abcdefghijklmnopqrstuvwxyz
%                        ABCDEFGHIJKLMNOPQRSTUVWXYZ!"
%                        #$%&\'()*+,-./:;<=>?@[\\]^_`{|}~ \t\n\r\x0b\x0c
\\
stats-of-text-length    & Computes the mean, median, mode, min, max of string lengths from a columns entries.\\
	\rowcolor{Gray}
stats-of-numerical-type & Computes the mean, median, mode, min, max of entries from a column of numerical type.\\

prop-instances-per-class-in-knearestneighbours         & This finds the k columns from the training set with most similar names.
It then produces a vector of class proportions from these k neighbours.
Parameters:
"num-neighbours": The number of neighbours to find (k).\\
	\rowcolor{Gray}
mean-character-cosine-similarity-from-class-examples   & For some test instance t, this computes the character distribution 
from this instance's entries.  It then produces a vector of the average
cosine similarity of t with training instances of each class.\\
min-editdistance-from-class-examples                   & For some test instance t, this produces a vector of the minimum column name
edit distance between t and all training instances of each class.
Parameters:
"cache-size": Size of cache for storing edit distances of words 
"max-comparisons-per-class": Max number of training instances to compare
against for each class.\\

	\rowcolor{Gray}
min-wordnet-jcn-distance-from-class-examples            & For some test instance t, this produces a vector of the minimum column name
JCN WS4J distance between t and all training instances of each class.
Parameters:
"cache-size": Size of cache for storing wordnet distances of words 
"max-comparisons-per-class": Max number of training instances to compare
against for each class.\\

min-wordnet-lin-distance-from-class-examples            & For some test instance t, this produces a vector of the minimum column name
LIN WS4J distance between t and all training instances of each class.
Parameters:
"cache-size": Size of cache for storing wordnet distances of words 
"max-comparisons-per-class": Max number of training instances to compare
against for each class.\\

\end{tabular}
\end{table}

\ignore{
The following feature extractors produce a vector value.
Some of these are computed together to reduce computation time.
Others are values that are computed for all classes.

@@

\end{enumerate}
}

\subsection{\minizinc{} Model \label{Ann:MZ}}
